---
title: "Take-Home Exercise 1"
author: "Joshua TING"
date: "April 12, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
  freeze: true
---

# **The Private Real Estate Market**

***Caveat***

The analysis presented here is based on the author's best knowledge and understanding of the data available at the time of analysis. Any property purchase or investment, should be consulted with a qualified real estate professional who can provide tailored guidance based on your specific circumstances and objectives. The author holds no liability of any decisions or purchases related.

# **Introduction**

80% of Singapore residents reside in public housing while 20% resides in private housing. In the private housing market, land ownership is divided into two parts: land and strata. For landed property, the plot of land belongs to the owner while in strata property, the plot of land is jointly owned by the legal owners in the same development ([The Business Times, 2022](https://www.businesstimes.com.sg/property/landed-home-prices-set-stay-firm-if-not-trend-upwards)). A detached house, commonly known as bungalow, semi-detached house and terrace house may possess either be landed or strata titled. Hence, this analysis intends to explore the unit price (PSF\$) of the property transactions based on the land ownership.

## **Database**

The [REALIS](https://www.ura.gov.sg/reis/index) (Real Estate Information System) is an interest-based system that gives subscribers a comprehensive and detailed access to Singapore's private real estate data (URA, n.d.). Each data set constitutes a quarter of the year. A total of five data sets (Quarter 1 to 4 of 2023 and Quarter 1 of 2024) were retrieved from REALIS and will be used for the analysis.

# **Data Preparation**

## 1. Load Package

The following functions will be loaded using pacman:p_load() in R Packages.

```{r}
pacman::p_load(tidyverse,
               ggrepel, ggthemes,
               ggridges, ggdist, ggpubr,
               patchwork, cowplot, reshape2)
```

## 2. Import Data

Retrieving the data from REALIS, A total of five data sets (Quarter 1 to 4 of 2023 and Quarter 1 of 2024) will be used. The data set ranging from Quarter 1 to 4 of 2023 will be scaffolding the analysis while Quarter 1 of 2024 data set will be the focus of the analysis.

```{r}
q1data <- read_csv("data/ResidentialTransaction20240414220633.csv")
q2data <- read_csv("data/ResidentialTransaction20240308161109.csv")
q3data <- read_csv("data/ResidentialTransaction20240308161009.csv")
q4data <- read_csv("data/ResidentialTransaction20240308160736.csv")
q5data <- read_csv("data/ResidentialTransaction20240308160536.csv")
```

## 3. Glimpse of Data

By employing the glimpse() function, we are able to visualise the total number of rows and columns in the respective data sets. Additionally, It indicates the categories and type of data.

```{r}
glimpse(q1data)
glimpse(q2data)
glimpse(q3data)
glimpse(q4data)
glimpse(q5data)
```

## 4. Check for Duplicates

The duplicated() function, embedded in the base package, revealed that there are no similar observations in each data set.

```{r}
duplicates_q1data <- q1data[duplicated(q1data), ]
duplicates_q2data <- q2data[duplicated(q2data), ]
duplicates_q3data <- q3data[duplicated(q3data), ]
duplicates_q4data <- q4data[duplicated(q4data), ]
duplicates_q5data <- q5data[duplicated(q5data), ]
```

## 5. Check for Missing Values

The colSums() function, embedded in the base package, revealed that all data sets except for q3data has no missing values.

```{r}
colSums(is.na(q1data))
colSums(is.na(q2data))
colSums(is.na(q3data))
colSums(is.na(q4data))
colSums(is.na(q5data))
```

In q3data, it was revealed the column of "Area (SQM)" has a total of 6 rows (observations) with NA values. Therefore, the na.omit() function in the stats package was used to remove the missing data in q3data. The preceding step using the same function colSums() was executed in ensuring the missing data has been removed.

```{r}
q3data = q3data %>%
  na.omit()
colSums(is.na(q3data))
```

## 6. Calculating Average of "Unit Price (\$PSF)" for each data set

The average unit price (\$PSF) will be calculated in each quarter to facilitate the exploratory data analysis in the later section.

```{r}
q1avg <- q1data %>% summarize(avg_unit_price = mean(`Unit Price ($ PSF)`))
q2avg <- q2data %>% summarize(avg_unit_price = mean(`Unit Price ($ PSF)`))
q3avg <- q3data %>% summarize(avg_unit_price = mean(`Unit Price ($ PSF)`))
q4avg <- q4data %>% summarize(avg_unit_price = mean(`Unit Price ($ PSF)`))
q5avg <- q5data %>% summarize(avg_unit_price = mean(`Unit Price ($ PSF)`))
```

## 7. Combine the average values into a single data frame

Following which, the average unit price (\$PSF) will be populated into a data frame and subsequently, into a line graph. The function "levels" was used in ensuring the data is populated in accordance to the chronological order.

```{r}
avg_data <- data.frame(
  Quarter = factor(c("Q1 2023", "Q2 2023", "Q3 2023", "Q4 2023", "Q1 2024"), levels = c("Q1 2023", "Q2 2023", "Q3 2023", "Q4 2023", "Q1 2024")),
  avg_unit_price = c(q1avg$avg_unit_price, q2avg$avg_unit_price, q3avg$avg_unit_price, q4avg$avg_unit_price, q5avg$avg_unit_price)
)
```

## Summary of Data Preparation

The five data sets were relatively clean and approriate for the analysis. Necessary steps were taken in ensuring the data is cleaned such as checking for duplicates and missing values. The following section will the essence of this analysis.

# **Exploratory Data Analysis**

Adhering to the aim of this analysis, the first observation will looked into the average unit price (\$PSF) of five quarters; quarter 1 of 2023 to quarter 1 of 2024 while the second observation will focus on quarter 1 of 2024.

### Observation 1: {.tabset}

Based on the line graph from Q1 2023 to Q1 2024, there is a spike of the Average Unit Price (\$PSF) in Q2 2023, approximately 4.5% from Q1 2023. Thereafter, the price normalised to the range of \$1825 - \$1850 and subsequently dipped to \$1812 in Q1 2024. The spike in Q2 2023 might have been attributed to the increased demand of private housing prior to the increase of Additional Buyer's Stamp Duty (ABSD) starting 27th April 2023 as part of the Government's cooling measures of the property measures ([CNA, 2023](https://www.channelnewsasia.com/singapore/property-cooling-measures-absd-additional-buyers-stamp-duty-doubled-foreigners-increase-3446376)). The aftereffect of the cooling measures can be visualised in the dipped of the Average Unit Price (\$PSF) from Q3 2023 to Q1 2024. This aligns with the published article by URA indicating the price and sales momentum in the private residential market were sluggish ([URA, 2024](https://www.ura.gov.sg/Corporate/Media-Room/Media-Releases/pr24-11)).

::: panel-tabset
## Code

```{r eval=FALSE}
ggplot(avg_data, aes(x = Quarter, y = avg_unit_price)) +
  geom_line(group=1) + 
  geom_point() + 
  geom_text(aes(label = paste("$",round(avg_unit_price))), vjust = 0.5, hjust = +1.3, colour="blue") +
  labs(
    title = element_text("Average Unit Price ($ PSF) Over 5 Quarters"),
    x = "Quarter of The Year",
    y = "Unit Price ($ PSF)"
  ) +
  theme(plot.title=element_text(hjust = 0.5, face = "bold")) +
  theme_gray()
```

## Output

```{r echo==FALSE}
ggplot(avg_data, aes(x = Quarter, y = avg_unit_price)) +
  geom_line(group=1) + 
  geom_point() + 
  geom_text(aes(label = paste("$",round(avg_unit_price))), vjust = 0.5, hjust = +1.3, colour="blue") +
  labs(
    title = element_text("Average Unit Price ($ PSF) Over 5 Quarters"),
    x = "Quarter of The Year",
    y = "Unit Price ($ PSF)"
  ) +
  theme(plot.title=element_text(hjust = 0.5, face = "bold")) +
  theme_gray()
```
:::

### Observation 2:

Q1 2024 was chosen to be examine further due to the sluggish private real estate transactions.

In a bird's eye view, it can be seen that the average PSF of the landed type is higher than the strata type with reference to Terrace House and Semi-Detached House. Additionally, the volume of transactions are significantly more in land than strata. However, the outliers, seen in the right-tail of the strata type, is much more than the land type. This might indicate despite the demand for landed-related transactions, there are still buyers that are willing to pay higher unit prices (psf) of strata-titled properties.

::: panel-tabset
## Code

```{r eval=FALSE}
combined_data <- q5data

ridgeline_combined <- ggplot(combined_data, aes(x = `Unit Price ($ PSF)`, y = `Property Type`, fill = `Property Type`)) +
  geom_density_ridges(alpha = 0.5) +
  scale_fill_viridis_d() +
  theme_ridges() +
  labs(title = "Ridgeline Plot for Strata and Landed in Q1 2024",
       x = "Unit Price ($ PSF)", y = NULL) +
  facet_wrap(~`Type of Area`, ncol = 1, labeller = labeller(`Type of Area` = c("Strata", "Land"))) +
  theme(legend.text = element_text(size = 5)) 

print(ridgeline_combined)
```

## Output

```{r echo=FALSE}
combined_data <- q5data

ridgeline_combined <- ggplot(combined_data, aes(x = `Unit Price ($ PSF)`, y = `Property Type`, fill = `Property Type`)) +
  geom_density_ridges(alpha = 0.5) +
  scale_fill_viridis_d() +
  theme_ridges() +
  labs(title = "Ridgeline Plot for Strata and Landed in Q1 2024",
       x = "Unit Price ($ PSF)", y = NULL) +
  facet_wrap(~`Type of Area`, ncol = 1, labeller = labeller(`Type of Area` = c("Strata", "Land"))) +
  theme(legend.text = element_text(size = 5)) 

print(ridgeline_combined)
```
:::

### Limitations:

1.  Only 5 data sets (5 quarters) were used thus unable to display a clear fluctuation of the average PSF prices over a spectrum of years

2.  Analysis did not explore transaction prices and its relation with districts or regions (CCR, RCR, OCR)

3.  Analysis did not look at transaction price as a totality. Despite the data showing higher unit prices (PSF) for strata-titled properties, the transaction prices might be lower compared to the landed-titled properties due to the property area

### Suggestions:

1.  To include data sets from 2019 - 2022 in ensuring a comprehensive view of the prices in the private real estate market

2.  To explore transactions prices and its sensitivity towards regions in Singapore

# **References:**

CNA. (2023). Property Cooling Measures. <https://www.channelnewsasia.com/singapore/property-cooling-measures-absd-additional-buyers-stamp-duty-doubled-foreigners-increase-3446376>

The Business Times. (2022). Landed home prices set to stay firm, if not trend upwards. <https://www.businesstimes.com.sg/property/landed-home-prices-set-stay-firm-if-not-trend-upwards>

URA. (2023). REALIS. <https://www.ura.gov.sg/-/media/Corporate/Property/REALIS/User%20Guide/REALIS_eBrochure.pdf>

URA. (2024). <https://www.ura.gov.sg/Corporate/Media-Room/Media-Releases/pr24-11>
