---
title: "Take Home Exercise 3"
author: "Joshua TING"
date: "May 18, 2024"
date-modified: "last-modified"
format:
  html:
    code-fold: true
    code-summary: "Code Chunk"
    number-sections: true
execute:
  eval: true
  echo: true
  warning: false
  freeze: true
  message: false
editor: source
---

## 1. Introduction - Mini-Challenge 3

Setting the Scene

The business community in Oceanus is dynamic with new startups, mergers, acquisitions, and investments. FishEye International closely watches business records to keep tabs on commercial fishing operators. FishEye’s goal is to identify and prevent illegal fishing in the region’s sensitive marine ecosystem. Analysts are working with company records that show ownership, shareholders, transactions, and information about the typical products and services of each entity. FishEye’s analysts have a hybrid automated/manual process to transform the data into CatchNet: the Oceanus Knowledge Graph.

In the past year, Oceanus’s commercial fishing business community was rocked by the news that SouthSeafood Express Corp was caught fishing illegally. FishEye wants to understand temporal patterns and infer what may be happening in Oceanus’s fishing marketplace because of SouthSeafood Express Corp’s illegal behavior and eventual closure. The competitive nature of Oceanus’s fishing market may cause some businesses to react aggressively to capture SouthSeafood Express Corp’s business while other reactions may come from the awareness that illegal fishing does not go undetected and unpunished.

Question 1: FishEye analysts want to better visualize changes in corporate structures over time. Create a visual analytics approach that analysts can use to highlight temporal patterns and changes in corporate structures. Examine the most active people and businesses using visual analytics.

Question 4: Identify the network associated with SouthSeafood Express Corp and visualize how this network and competing businesses change as a result of their illegal fishing behavior. Which companies benefited from SouthSeafood Express Corp legal troubles? Are there other suspicious transactions that may be related to illegal fishing? Provide visual evidence for your conclusions.

## 1. Loading Packages

```{r}
pacman::p_load(jsonlite, tidygraph, ggraph, kableExtra,
               visNetwork, graphlayouts, ggforce, textstem, 
               skimr, tidytext, tidyverse, gganimate,dplyr, lubridate, DT)
```

## 2. Data Preparation

## Data Import

In the code chunk below, `fromJSON()` of **jsonlite** package is used to import *MC3.json* into R environment.

```{r}
mc3_data <- fromJSON("data/MC3_/mc3.json")
```

### **Extracting edges**

The code chunk below will be used to extract the *links* data.frame of *mc3_data* and save it as a tibble data.frame called *mc3_edges*.

```{r}
mc3_edges <- as_tibble(mc3_data$links) %>% 
  unnest(source) %>% 
  distinct() %>% 
  mutate(source = as.character(source),
         target = as.character(target),
         type = as.character(type),
         startdate = as_datetime(start_date)) %>%
  group_by(source, target, type, startdate) %>% 
  summarise(weights = n()) %>% 
  filter(source != target) %>%
  ungroup()

head(mc3_edges)
```

Splitting Words

```{r}
word_list <- strsplit(mc3_edges$type, "\\.")
```

```{r}
max_elements <- max(lengths(word_list))
```

```{r}
word_list_padded <- lapply(word_list, function(x) c(x, rep(NA, max_elements - length(x))))
```

```{r}
word_df <- do.call(rbind, word_list_padded)
colnames(word_df) <- paste0("event", 1:max_elements)
```

```{r}
word_df <- as_tibble(word_df) %>%
  select(event2, event3)
class(word_df)
```

```{r}
mc3_edges <- mc3_edges %>%
  cbind(word_df)
```

::: callout-note
Learning from Code Chunk Above

-   `distinct()` is used to ensure that there will be no duplicated records.

-   `mutate()` and `as.character()` are used to convert the field data type from list to character.

-   `group_by()` and `summarise()` are used to count the number of unique links.

-   the `filter(source!=target)` is to ensure that no record with similar source and target.
:::

### **Extracting nodes**

The code chunk below will be used to extract the *nodes* data.frame of *mc3_data* and save it as a tibble data.frame called *mc3_nodes*.

```{r}
# extract all nodes from graph
mc3_nodes <- as_tibble(mc3_data$nodes) %>% 
  mutate(country = as.character(country),
         id = as.character(id),
         revenue = as.numeric(as.character(revenue)),
         type = as.character(type)) %>%
  select(id, country, type, revenue)

# extract all nodes from edges
id1 <- mc3_edges %>%
  select(source, type) %>%
  rename(id = source) %>% 
  mutate(country = NA, revenue = NA) %>% 
  select(id, country, type, revenue)

id2 <- mc3_edges %>%
  select(target, type) %>%
  rename(id = target) %>% 
  mutate(country = NA, revenue = NA) %>% 
  select(id, country, type, revenue)

additional_nodes <- rbind(id1, id2) %>% 
  distinct %>% 
  filter(!id %in% mc3_nodes[["id"]])

# combine all nodes
mc3_nodes_updated <- rbind(mc3_nodes, additional_nodes) %>%
  distinct()

head(mc3_nodes_updated)
```

::: callout-note
Learning From Code Chunk Above

-   `mutate()` and `as.character()` are used to convert the field data type from list to character.

-   To convert *revenue_omu* from list data type to numeric data type, we need to convert the values into character first by using `as.character()`. Then, `as.numeric()` will be used to convert them into numeric data type.

-   `select()` is used to re-organise the order of the fields.
:::

## **3. Data Exploration**

### **Exploring Edges DF**

In the code chunk below, [`skim()`](https://docs.ropensci.org/skimr/reference/skim.html) of [**skimr**](https://docs.ropensci.org/skimr/) package is used to display the summary statistics of *mc3_edges* tibble data frame.

```{r}
skim(mc3_edges)
```

In the code chunk below, `datatable()` of DT package is used to display mc3_edges tibble data frame as an interactive table on the html document.

```{r}
DT::datatable(mc3_edges)
```

```{r}
ggplot(data = mc3_edges,
       aes(x = type)) +
  geom_bar()
```

In the above graph, it shows the counts of people that either is a beneficial owner, shareholder, employee or family member of the organisation. A substantial of the graph belongs to the shareholders.

## **Initial Network Visualisation and Analysis**

### **Building network model with tidygraph**

```{r}
mc3_graph <- tbl_graph(nodes = mc3_nodes_updated,
                       edges = mc3_edges,
                       directed = FALSE) %>%
  mutate(betweenness_centrality = centrality_betweenness(),
         closeness_centrality = centrality_closeness())
```

5.1 Tidying Nodes

```{r}
mc3_nodes1 <- mc3_nodes_updated %>%
  mutate(type = gsub("[(]", "", type)) %>% 
  mutate(type = gsub("\"", "", type)) %>%
  mutate(type = gsub("[)]", "", type)) 
```


```{r}
graph <- tbl_graph(nodes = mc3_nodes_updated, edges = mc3_edges, directed = TRUE)
```


```{r}
period1_start <- as_date("2023-01-01")
period1_end <- as_date("2023-06-30")
period2_start <- as_date("2023-07-01")
period2_end <- as_date("2023-12-31")

```

```{r}

# Create graph objects for each period
graph_period1 <- tbl_graph(nodes = mc3_nodes_updated, edges = mc3_edges, directed = TRUE)
graph_period2 <- tbl_graph(nodes = mc3_nodes_updated, edges = mc3_edges, directed = TRUE)
```

```{r}
# Plot for period 1
plot1 <- ggraph(mc3_graph, layout = 'fr') +
  geom_edge_link() +
  geom_node_point() +
  theme_graph() +
  ggtitle("Plot 1")
```


```{r}
set.seed(123)
mc3_graph %>%
  filter(betweenness_centrality >= 1000000) %>%
ggraph(layout = "fr") +
  geom_edge_link(aes(#width= weights,
                     alpha=0.5)) +
  geom_node_point(aes(
    size = betweenness_centrality,
    color = type,
    alpha = 0.3)) +
  geom_node_label(aes(label = id),repel=TRUE, size=2.5, alpha = 0.8) +
  scale_size_continuous(range=c(1,10)) +
  theme_graph() +
  labs(title = 'Initial network visualisation',
       subtitle = 'Entities with betweenness scores > 1,000,000')
```

