---
title: "In Class Exercise 1"
author: "Joshua TING"
date: "April 13, 2024"
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting Started

## Loading R Packages

```{r}
pacman::p_load(tidyverse)
```

::: callout-note
## p_load

The code chunk above uses **'p_load()'** of the pacman package to load the [tidyverse](https://www.tidyverse.org/packages/) family of packages. It ensures that the package is installed.
:::

## Importing Data

```{r}
realis <- read_csv("data/realis2019.csv")
```

::: callout-note
## readr

The code chunk above imports *realis.2019.csv* into the R envrionment by using [read.csv()](https://readr.tidyverse.org/reference/read_delim.html) of the [readr()](https://lnealicia-isss608.netlify.app/in-class_ex/in-class_ex01/in-class_ex01) package.

Readr is part of the tidyverse package.
:::

## ggpplot

```{r}
ggplot(data=realis, 
       aes(x = `Unit Price ($ psm)`)) +
  geom_histogram() +
  ggtitle("Distribution of Unit Price ($ psm)")
```

```{r}
glimpse(realis)
```

::: callout-note
## ggplot

[ggplot2](https://ggplot2.tidyverse.org/) is an R package for creating data-driven graphics based on [***The Grammar of Graphics***](https://www.amazon.com/Grammar-Graphics-Statistics-Computing/dp/0387245448/ref=as_li_ss_tl). It is also part of the tidyverse family specially designed for visual exploration and communication.
:::
